---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Study Group 4"
date: "1/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive vocal patterns in schizophrenia? Report how many papers report quantitative estimates, comment on what percentage of the overall studies reviewed they represent (see PRISMA chart) your method to analyze them, the estimated effect size of the difference (mean effect size and standard error) and forest plots representing it. N.B. Only measures of pitch mean and pitch sd are required for the assignment (all the other measures have been removed from the dataset for the sake of simplicity). 

2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

## Tips on the process to follow:

- Download the data on all published articles analyzing voice in schizophrenia and the prisma chart as reference of all articles found and reviewed
Data: https://www.dropbox.com/s/0l9ur0gaabr80a8/Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx?dl=0
Prisma chart: https://www.dropbox.com/s/vbjf6ff73dv9tru/PRISMA%202009%20flow%20diagram_PP.pptx?dl=0 
- Look through the dataset to find out which columns to use, and if there is any additional information written as comments (real world data is always messy!).
    * Hint: PITCH_F0M and PITCH_F0SD group of variables are what you need
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2

```{r}
#Kristine
#Loading packages
pacman::p_load(tidyverse, metafor, readxl, lme4)

#Loading data 
meta <- read_excel("Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx")

#cleaning data 
sub <- meta %>% select(StudyID:DIAGNOSIS, PITCH_F0_HC_M, PITCH_F0_SZ_M, PITCH_F0_HC_SD, PITCH_F0_SZ_SD, -Specification)

  #Sarahs function
delete.na <- function(DF, n=0) {
  DF[rowSums(is.na(DF)) <= n,]
}
  #Deleting studies that we are not interested in
sub <- delete.na(sub, 3)

  #Merging clean and subsetted data
meta_clean <- left_join(sub, meta)

  #Choosing columns we are interested in 
meta_clean <- meta_clean %>% select(StudyID:Article, SAMPLE_SIZE_HC:SAMPLE_SIZE_SZ, TYPE_OF_TASK, PITCH_F0_HC_M:PITCH_F0_SZ_SD)

```

#Meta-analysis
```{r}
#Nanna
#Calculating effect size
  #In the first step we use the escalc() function to calculate effect sizes and their variance.

met1 <-  escalc(measure = "SMD", n1i = SAMPLE_SIZE_HC, n2i = SAMPLE_SIZE_SZ, m1i = PITCH_F0_HC_M, m2i = PITCH_F0_SZ_M, 
    sd1i = PITCH_F0_HC_SD, sd2i = PITCH_F0_SZ_SD, data = meta_clean)

#Making models 

# Mixed effects implementation 
m <- lmer(yi ~ 1 + (1 | StudyID), weights = 1/vi, data= met1, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))

summary(m)

#Meta-analysis optimization: 
m2 <-rma (yi, vi, data = met1, slab = StudyID)

#Results 
summary(m2)
forest(m2)

#Influential studies
inf <- influence(m2)
print(inf)
plot(inf)
  #Seems like one study is an outliers 

#Funnel plot
funnel(m2)
  #Funnel plot looks okay
```
#Meta-analysis with data from assignment 3 
```{r}
#Sarah
#Loading data from assignment 3 
own <- read_csv("SchizophreniaData.csv")

#Cleaning data for controls
  #Make columns with mean and sd for control in study 1

control_1 <- subset(own, Diagnosis == "Control" & Study == 1)

hc1_m <- mean(control_1$pitch_mean)
hc1_sd <- sd(control_1$pitch_mean)

  #Make columns with mean and sd for control in study 2

control_2 <- subset(own, Diagnosis == "Control" & Study == 2)

hc2_m <- mean(control_2$pitch_mean)
hc2_sd <- sd(control_2$pitch_mean)

  #Make columns with mean and sd for control in study 3

control_3 <- subset(own, Diagnosis == "Control" & Study == 3)

hc3_m <- mean(control_3$pitch_mean)
hc3_sd <- sd(control_3$pitch_mean)

  #Make columns with mean and sd for control in study 4

control_4 <- subset(own, Diagnosis == "Control" & Study == 4)

hc4_m <- mean(control_4$pitch_mean)
hc4_sd <- sd(control_4$pitch_mean)

#Cleaning data for schizophrenia
  #Make columns with mean and sd for SZ for study 1

sz_1 <- subset(own, Diagnosis == "Schizophrenia" & Study == 1)
sz1_m <- mean(sz_1$pitch_mean)
sz1_sd <- sd(sz_1$pitch_mean)

  #Make columns with mean and sd for SZ for study 2
sz_2 <- subset(own, Diagnosis == "Schizophrenia" & Study == 2)
sz2_m <- mean(sz_2$pitch_mean)
sz2_sd <- sd(sz_2$pitch_mean)


  #Make columns with mean and sd for SZ for study 3
sz_3 <- subset(own, Diagnosis == "Schizophrenia" & Study == 3)
sz3_m <- mean(sz_3$pitch_mean)
sz3_sd <- sd(sz_3$pitch_mean)


  #Make columns with mean and sd for SZ for study 4
sz_4 <- subset(own, Diagnosis == "Schizophrenia" & Study == 4)
sz4_m <- mean(sz_4$pitch_mean)
sz4_sd <- sd(sz_4$pitch_mean)

```
```{r}
#Julie
#Making new dataframe with studies from assignment 3 

met2 <- met1 %>% add_row(StudyID = 51,   
                                      Title = "Voice patterns in schizophrenia: A systematic review and Bayesian meta-analysis",   
                                      Authors = "Parola, A., Simonsen, A., Bliksted, V., & Fusaroli, R.",    
                                      Year_publication = 2019,
                 Article = "Fusorali et al.",
                 SAMPLE_SIZE_HC = nrow(control_1),
                 SAMPLE_SIZE_SZ = nrow(sz_1),
                 TYPE_OF_TASK = NA,
                 PITCH_F0_HC_M = hc1_m,
                 PITCH_F0_HC_SD = hc1_sd,
                 PITCH_F0_SZ_M = sz1_m,
                 PITCH_F0_SZ_SD = sz1_sd,
                 yi =NA, 
                 vi = NA) 

met2 <- met2 %>% add_row(StudyID = 52,   
                                      Title = "Voice patterns in schizophrenia: A systematic review and Bayesian meta-analysis",   
                                      Authors = "Parola, A., Simonsen, A., Bliksted, V., & Fusaroli, R.",    
                                      Year_publication = 2019,
                 Article = "Fusorali et al.",
                 SAMPLE_SIZE_HC = nrow(control_2),
                 SAMPLE_SIZE_SZ = nrow(sz_2),
                 TYPE_OF_TASK = NA,
                 PITCH_F0_HC_M = hc2_m,
                 PITCH_F0_HC_SD = hc2_sd,
                 PITCH_F0_SZ_M = sz2_m,
                 PITCH_F0_SZ_SD = sz2_sd,
                 yi =NA, 
                 vi = NA) 

met2 <- met2 %>% add_row(StudyID = 53,   
                                      Title = "Voice patterns in schizophrenia: A systematic review and Bayesian meta-analysis",   
                                      Authors = "Parola, A., Simonsen, A., Bliksted, V., & Fusaroli, R.",    
                                      Year_publication = 2019,
                 Article = "Fusorali et al.",
                 SAMPLE_SIZE_HC = nrow(control_3),
                 SAMPLE_SIZE_SZ = nrow(sz_3),
                 TYPE_OF_TASK = NA,
                 PITCH_F0_HC_M = hc3_m,
                 PITCH_F0_HC_SD = hc3_sd,
                 PITCH_F0_SZ_M = sz3_m,
                 PITCH_F0_SZ_SD = sz3_sd,
                 yi =NA, 
                 vi = NA) 

met2 <- met2 %>% add_row(StudyID = 54,    
                                      Title = "Voice patterns in schizophrenia: A systematic review and Bayesian meta-analysis",   
                                      Authors = "Parola, A., Simonsen, A., Bliksted, V., & Fusaroli, R.",    
                                      Year_publication = 2019,
                 Article = "Fusorali et al.",
                 SAMPLE_SIZE_HC = nrow(control_4),
                 SAMPLE_SIZE_SZ = nrow(sz_4),
                 TYPE_OF_TASK = NA,
                 PITCH_F0_HC_M = hc4_m,
                 PITCH_F0_HC_SD = hc4_sd,
                 PITCH_F0_SZ_M = sz4_m,
                 PITCH_F0_SZ_SD = sz4_sd,
                 yi =NA, 
                 vi = NA)
```

```{r}
#Martine
#Calculating effect size 

new_met <- escalc(measure = "SMD", n1i = SAMPLE_SIZE_HC, n2i = SAMPLE_SIZE_SZ, m1i = PITCH_F0_HC_M, m2i = PITCH_F0_SZ_M, 
    sd1i = PITCH_F0_HC_SD, sd2i = PITCH_F0_SZ_SD, data = met2)

#Mixed effects implementation 

m_tot <- lmer(yi ~ 1 + (1 | StudyID), weights = 1/vi, data= new_met, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))

summary(m_tot)

#Meta-analysis optimization: 
m_tot2 <-rma (yi, vi, data = new_met, slab = StudyID)

#Results 
summary(m_tot2)
forest(m_tot2)

#Influential studies
inf_tot <- influence(m_tot2)
print(inf_tot)
plot(inf_tot)

#Funnel plot
funnel(m_tot2)
```


